---
title: "EyetrackingDataCleanUp"
author: "RF"
date: "2/1/2018"
output: html_document
---

# Here we load libraries and data (from 2018)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(readr,groupdata2,ggplot2,tidyverse)

# Both experiments are in the same files
setwd('~/Dropbox (Personal)/My courses/2018 - Eye tracking')
Samples=read_delim('eyetracking_2018_samples.txt',delim='\t')
Fixations=read_delim('eyetracking_2018_fixations.txt',delim='\t')
Saccades=read_delim('eyetracking_2018_saccades.txt',delim='\t')
```

# Here we clean up the data
- We only select useful columns
- We rename the columns in more concise and meaningful ways
- We add the task (Visual Search vs. Social Engagement)
- We add time as counted from the start of each trial
- We check whether the trials are of the length we expect (6s for videos, 10s/20s/40s for images)

```{r}

AddTask <- function(df){
  for (s in unique(df$ParticipantID)){
    if (max(df$Trial[df$ParticipantID==s])==8){
      df$Task[df$ParticipantID==s]='SocialEngagement'
    } else {
      df$Task[df$ParticipantID==s]='VisualSearch'
    }
  }
  df$ParticipantGender[grepl("f|F",df$ParticipantID)]='female'
  df$ParticipantGender[grepl("m|M",df$ParticipantID)]='male'
  
  df$SearchOrder[grepl("f1|F1",df$ParticipantID)]=1
  df$SearchOrder[grepl("m1|M1",df$ParticipantID)]=1
  df$SearchOrder[grepl("f2|F2",df$ParticipantID)]=2
  df$SearchOrder[grepl("m2|M2",df$ParticipantID)]=2
  
  return(df)
}

Fixations = Fixations %>% 
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    Fixation=CURRENT_FIX_INDEX,
    Duration=CURRENT_FIX_DURATION,
    PupilSize=CURRENT_FIX_PUPIL,
    PositionX=CURRENT_FIX_X,
    PositionY=CURRENT_FIX_Y,
    StartTime=CURRENT_FIX_START,
    EndTime=CURRENT_FIX_END)


Fixations=AddTask(Fixations)


Saccades = Saccades %>%
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    Saccade=CURRENT_SAC_INDEX,
    Direction=CURRENT_SAC_DIRECTION,
    Amplitude=CURRENT_SAC_AMPLITUDE,
    Duration=CURRENT_SAC_DURATION,
    PeakVelocity=CURRENT_SAC_PEAK_VELOCITY,
    MeanVelocity=CURRENT_SAC_AVG_VELOCITY,
    Blink=CURRENT_SAC_CONTAINS_BLINK,
    StartTime=CURRENT_SAC_START_TIME,
    EndTime=CURRENT_SAC_END_TIME,
    StartX=CURRENT_SAC_START_X,
    StartY=CURRENT_SAC_START_Y,
    EndX=CURRENT_SAC_END_X,
    EndY=CURRENT_SAC_END_Y
  )
Saccades$Amplitude=as.numeric(Saccades$Amplitude)
Saccades$MeanVelocity=as.numeric(Saccades$MeanVelocity)
Saccades$EndX=as.numeric(Saccades$EndX)
Saccades$EndY=as.numeric(Saccades$EndY)

Saccades=AddTask(Saccades)

Samples = Samples %>%
  select(
    ParticipantID=RECORDING_SESSION_LABEL,
    Trial=TRIAL_INDEX,
    TimeStamp=TIMESTAMP,
    PositionX=RIGHT_GAZE_X,
    PositionY=RIGHT_GAZE_Y,
    PupilSize=RIGHT_PUPIL_SIZE,
    Saccade=RIGHT_IN_SACCADE,
    Blink=RIGHT_IN_BLINK,
    MeanVelocityX=RIGHT_VELOCITY_X,
    MeanVelocityY=RIGHT_VELOCITY_Y,
    MeanAccellerationX=RIGHT_ACCELLERATION_X,
    MeanAccellerationY=RIGHT_ACCELLERATION_Y
  )

Samples$MeanVelocityY=as.numeric(Samples$MeanVelocityY)
Samples$MeanVelocityX=as.numeric(Samples$MeanVelocityX)
Samples$MeanAccellerationY=as.numeric(Samples$MeanAccellerationY)
Samples$MeanAccellerationX=as.numeric(Samples$MeanAccellerationX)

Samples=AddTask(Samples)

# Add time from trial start

Samples <- Samples %>%
  group_by(ParticipantID,Trial) %>%
  mutate(TrialTime = TimeStamp - min(TimeStamp))

## Double check trial timing. It's all in place!
X <- Samples %>%
   group_by(ParticipantID,Trial) %>%
   dplyr::summarize(LengthSeconds = max(TrialTime)/1000)
print(X)
```

## Import fixation number and saccade number from processed data to samples
Since the samples do not exclude artifacts and do not contain fixation and saccade number, we import this data from Fixations and Saccades datasets

```{r}

for (t in seq(nrow(Fixations))){
  StartTime=Fixations$StartTime[t]
  EndTime=Fixations$EndTime[t]
  ID=Fixations$ParticipantID[t]
  Trial=Fixations$Trial[t]
  FixN=Fixations$Fixation[t]
  Samples$Fixation[Samples$TrialTime >= StartTime  & Samples$TrialTime <= EndTime & Samples$ParticipantID==ID & Samples$Trial==Trial]=FixN
}


for (t in seq(nrow(Saccades))){
  StartTime=Saccades$StartTime[t]
  EndTime=Saccades$EndTime[t]
  ID=Saccades$ParticipantID[t]
  Trial=Saccades$Trial[t]
  SacN=Saccades$Fixation[t]
  Samples$SaccadeN[Samples$TrialTime >= StartTime  & Samples$TrialTime <= EndTime & Samples$ParticipantID==ID & Samples$Trial==Trial]=SacN
}

write_csv(Saccades,"SaccadesV1.csv")
write_csv(Fixations,"FixationsV1.csv")
write_csv(Samples,"SamplesV1.csv")

```

## Merge with data about experimental conditions
We merge the data with the experimental logs.

```{r}

Fixations=read_csv('FixationsV1.csv')
Saccades=read_csv('SaccadesV1.csv')
Samples=read_csv('SamplesV1.csv')

## THEN GO TO THE LOG FILES AND MATCH the data from the logfiles to the samples/fixations/saccades (to have conditions, and so forth)
## ADD GROUP PARTICIPANT AND GENDER TO THE TABLES

Order1<- read_tsv('ForagingOrder1.txt',col_names = F)
Order1$Trial=1:10
Order1$SearchOrder=1
Order1$SearchType=c("Search","Search","Search","Search","Search","Count","Count","Count","Count","Count")
Order2<- read_tsv('ForagingOrder1.txt',col_names = F)
Order2$Trial=1:10
Order2$SearchOrder=2
Order2$SearchType=c("Count","Count","Count","Count","Count","Search","Search","Search","Search","Search")
Order=rbind(Order1,Order2)
Order=Order %>% rename(Stimulus=X1)

Saccades=merge(Saccades,Order,all=T)
Fixations=merge(Fixations,Order,all=T)
Samples=merge(Samples,Order,all=T)

setwd('PupilsLogs')
fileList<-list.files(".",pattern="csv")
logs=NULL
for (f in fileList){
  d<-read_csv(f)
  d=d %>% rename(Trial=X1,ParticipantID=subject)
  d$Trial=d$Trial+1
  for (n in seq(nrow(d))){
    x<-str_split(d$video[n],"_")[[1]]
    d$ActorGender[n]=x[1]
    d$Directionality[n]=x[4]
    d$Ostension[n]=x[5]
  }
  if (exists('logs')){logs=rbind(logs,d)} else {logs=d}
}

Saccades=merge(Saccades,logs,all=T)
Saccades$Stimulus[Saccades$Task=='SocialEngagement']=Saccades$video[Saccades$Task=='SocialEngagement']
Saccades$video=NULL
Fixations=merge(Fixations,logs,all=T)
Fixations=Fixations[!is.na(Fixations$Task),]
Fixations$Stimulus[Fixations$Task=='SocialEngagement']=Fixations$video[Fixations$Task=='SocialEngagement']
Fixations$video=NULL
Samples=merge(Samples,logs,all=T)
Samples$Stimulus[Samples$Task=='SocialEngagement']=Samples$video[Samples$Task=='SocialEngagement']
Samples$video=NULL
setwd('..')
write_csv(Saccades,"SaccadesV2.csv")
write_csv(Fixations,"FixationsV2.csv")
write_csv(Samples,"SamplesV2.csv")
```

# We now integrate the datasets with all the available data from 2017

```{r}

Fixations=read_csv('FixationsV2.csv')
Saccades=read_csv('SaccadesV2.csv')
Samples=read_csv('SamplesV2.csv')

ForagingFixations <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/FORAGING_FixationsClean.csv")
ForagingFixations$Task='VisualSearch'
ForagingSaccades <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/FORAGING_SaccadesClean.csv")
ForagingSaccades$Task='VisualSearch'
PupilsFixations <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/PUPIL_FixationsClean.csv")
PupilsFixations$Task='SocialEngagement'
PupilsSamples <- read_csv("~/Dropbox (Personal)/My courses/old/2017 - EyeTracking/Data/PUPIL_SamplesClean.csv")
PupilsSamples$Task='SocialEngagement'


ForagingFixations <- ForagingFixations %>% rename(
  SearchType=Condition,
  SearchOrder=Order,
  Duration=Duration,
  StartTime=CURRENT_FIX_START,
  EndTime=CURRENT_FIX_END,
  PositionX=CURRENT_FIX_X,
  PositionY=CURRENT_FIX_Y
)

ForagingFixations[,c('PupilSize','ActorGender','Directionality','Ostension','ParticipantGender')]=NA

PupilsFixations <- PupilsFixations %>% rename(
  ParticipantID=participant_id,
  Trial=trial,
  Fixation=current_fix_index,
  Duration=current_fix_duration,
  StartTime=current_fix_start,
  EndTime=current_fix_end,
  PositionX=current_fix_x,
  PositionY=current_fix_y,
  Task=Task,
  PupilSize=current_fix_pupil,
  Stimulus=video,
  ActorGender=gender,
  Directionality=directed,
  Ostension=ostension
)

PupilsFixations[,c('Condition','SearchType','SearchOrder','ParticipantGender')]=NA

Fixations=Fixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]

ForagingFixations=ForagingFixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]

PupilsFixations=PupilsFixations[,c("ParticipantID","Trial","Fixation","Duration","PupilSize","PositionX","PositionY","StartTime","EndTime","Task","SearchType","SearchOrder","Stimulus","Directionality","Ostension","ActorGender","ParticipantGender")]


Fixations2017=rbind(ForagingFixations,PupilsFixations)
Fixations2017$Year='2017'
Fixations$Year='2018'
Fixations=rbind(Fixations,Fixations2017)

ForagingSaccades <- ForagingSaccades %>% rename(
  SearchType=Condition,
  SearchOrder=Order,
  Direction=CURRENT_SAC_DIRECTION,
  Amplitude=CURRENT_SAC_AMPLITUDE,
  PeakVelocity=CURRENT_SAC_PEAK_VELOCITY,
  MeanVelocity=CURRENT_SAC_AVG_VELOCITY,
  Blink=CURRENT_SAC_CONTAINS_BLINK,
  StartTime=CURRENT_SAC_START_TIME,
  EndTime=CURRENT_SAC_END_TIME,
  StartX=CURRENT_SAC_START_X,
  StartY=CURRENT_SAC_START_Y,
  EndX=CURRENT_SAC_END_X,
  EndY=CURRENT_SAC_END_Y
)

ForagingSaccades[,c("ParticipantGender","Stimulus","ActorGender","Directionality","Ostension" )]=NA
ForagingSaccades$Year='2017'
Saccades=Saccades[,c("ParticipantID","Trial","Saccade","Direction","Amplitude","PeakVelocity","MeanVelocity","StartTime","EndTime","Duration", "StartX", "StartY","EndX","EndY","Blink","SearchType","SearchOrder","Task","ParticipantGender","Stimulus", "ActorGender" , "Directionality","Ostension")]
Saccades$Year='2018'  
Saccades=rbind(Saccades,ForagingSaccades)

PupilsSamples <- PupilsSamples %>% rename(
  ParticipantID=participant_id,
  Trial=trial,
  PupilSize=right_pupil_size,
  Fixation=fix_nr,
  Saccade=right_in_saccade,
  Blink=right_in_blink,
  ActorGender=gender,
  Directionality=directed,
  Ostension=ostension,
  Stimulus=video,
  TrialTime=task_time,
  Task=Task
)

PupilsSamples[,c("SearchOrder","TimeStamp","PositionX","PositionY","MeanVelocityX","MeanVelocityY","MeanAccellerationX","MeanAccellerationY","ParticipantGender","SearchType")]=NA
PupilsSamples$fixation=NULL

Samples=Samples[,colnames(PupilsSamples)] 
PupilsSamples$Year='2017'
Samples$Year='2018'
Samples=rbind(Samples,PupilsSamples)

Fixations$Ostension[Fixations$Ostension=='+o']='Ostensive'
Fixations$Ostension[Fixations$Ostension=='-o']='Private'
Samples$Ostension[Samples$Ostension=='+o']='Ostensive'
Samples$Ostension[Samples$Ostension=='-o']='Private'

Fixations$Directionality[Fixations$Directionality=='div']='Towards3rd'
Fixations$Directionality[Fixations$Directionality=='dir']='TowardsYou'
Samples$Directionality[Samples$Directionality=='div']='Towards3rd'
Samples$Directionality[Samples$Directionality=='dir']='TowardsYou'

Fixations$ActorGender[Fixations$ActorGender=='f']='Female'
Fixations$ActorGender[Fixations$ActorGender=='m']='Male'
Samples$ActorGender[Samples$ActorGender=='f']='Female'
Samples$ActorGender[Samples$ActorGender=='m']='Male'

Fixations$SearchOrder[Fixations$SearchOrder=='First']='1'
Fixations$SearchOrder[Fixations$SearchOrder=='Second']='2'
Saccades$SearchOrder[Saccades$SearchOrder=='First']='1'
Saccades$SearchOrder[Saccades$SearchOrder=='Second']='2'

write_csv(Saccades,"SaccadesV3.csv")
write_csv(Fixations,"FixationsV3.csv")
write_csv(Samples,"SamplesV3.csv")

```

